Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-03-01 ms.tex
brak

To first order our predicted average values of $\langle L_{X}/L_{\rm edd}\rangle$ increases only weakly with black hole mass.

for the average galactic star formation histories overlap with the data reasonably well.  Our prediction that $\langle L_{X}/L_{\rm edd}\rangle$ should increase with black hole mass appears to be in disagreement with observed decrease of $\langle L_{X}/L_{\rm edd}\rangle$ with $M_{\bullet}$ for both core and cusp galaxies.  However, better agreement is obtained by taking into account the fact that core(cusp) galaxie are more commonly associated with larger(smaller) black hole masses, such that when both samples are combined one predicts a flatter dependence of $\langle L_X/L_{\rm edd}\rangle$ of $M_{\bullet}$.   The large scatter observed in $L_{X}/L_{\rm edd}$ of typically 2-4 orders of magnitude can also be understood within a steady-accretion scenario due to the strong sensitivity of the accretion rate on the efficacy of gas heating, $\dot{M}/\dot{M}_{\rm edd} \propto v_{w}^{-3.8(-2.4)}$ for core(cusp) galaxies, respectively (Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]) and the equally sensitive dependence of the heating rate on stochastic intermittency in the star formation history (Fig.~NICK).

In the case of an old stellar population which produces insufficient heating to support steady accretion, this large scatter can in principle also be understood by the wide range of gas masses experienced over the course of a cyclic episode between instabilities and periodic nuclear outbursts (e.g.~\citealt{Ciotti+10}).  The cyclic model also predicts that high $L_{X}$ nuclei should be more common in galaxies with younger stellar ages.  This appears in tension with the finding by \citet{Pellegrini10} that galaxies with younger stellar populations are observed to host less X-ray luminous nuclei, at least for stellar ages $<3$ Gyr.  This observation can also be understood in the stable accretion picture by the fact that young stars provide sufficient stellar wind heating to produce a thermally stable accretion state.

However, the observed values of 
$\langle L_{X}/L_{\rm edd}\rangle$ do imply the accretion rates of
observed galaxies are below the maximum ones for thermally stable
accretion if $\epsilon_X \lesssim 10^{-4}$.  This suggests that most
nearby galaxies are consistent with accreting in a thermally stable
steady state, even if our estimate of their predicted heating does not
match the observed trend with $M_{\bullet}$ perfectly.  On the other
hand if $\epsilon_{X} \sim 10^{-5}$ then the trend would be consistent
with the accretion rate set by Ia blow-out.  Also note that in the
case of $\dot{M}-$dependent radiative efficiency ($\epsilon_X =
\dot{M}/\dot{M}_{\rm edd}$), the predicted branches of $\langle
L_{X}/L_{\rm edd}\rangle$ decreases with decreasing $M_{\bullet}$,
calling into question attempts to extrapolate the $\langle L_{X}
\rangle - M_{\bullet}$ relationship to low $M_{\bullet}$ in order to
constrain the occupation fraction of black holes in the lowest mass
galaxies (\citealt{Miller+15}).

Past studies have found that core galaxies show a wide range of nuclear radio and X-ray luminosities, while cusp galaxies are confined below a threshold (e.g., \citealt{Bender+89, Pellegrini99, Capetti&Balmaverde05}).

@@

Figure \ref{fig:bh_growth} shows that across all black hole masses, Compton heating alone cannot prevent thermal instabilities ($\dot{M}_{\rm C} > \dot{M}_{\rm TI}$), as will described more below.

>>>>

If heating in the nuclei of elliptical galaxies is characterized by that of the average star formation history of galaxies with similar mass, then to first order we predict that $\langle L_{X}/L_{\rm edd}\rangle$ should increase with black hole mass, though only weakly for $M_{\bullet} \lesssim 3\times 10^{8}M_{\odot}$.  This prediction is in tension with the observed trend that average Eddington ratio decreases (also weakly) with $M_{\bullet}$.  However, the physics setting $L_X$ based on the large-scale accretion rate is notoriously uncertain and there could exist black hole mass-dependent efficiency factors we are not taking into account.  For instance, if disk disk outflows carry away a substantial fraction of the infalling mass before it reaches sufficiently small radii to produce nuclear X-rays (e.g.~\citealt{Blandford&Begelman99}), then the angular momentum of the infalling gas (and hence that of stars in the galactic nucleus) might also affect the X-ray luminosity, potentially in a way that would depend systematically on the black hole mass.

It is also important to ask whether the steady-state accretion picture can explain the large large scatter (typically 2-3 orders of magnitude) observed in $L_{X}/L_{\rm edd}$.  Large scatter can in fact result from two effects: (1) the order of magnitude difference between $\dot{M}$ for core and cusp galaxies at fixed $v_w$ and $M_{\bullet}$ results in scatter due to the scatter in the Nuker stellar density profile $\Gamma$; (2) the strong sensitivity of the accretion rate on the gas heating rate, $\dot{M}/\dot{M}_{\rm edd} \propto v_{w}^{-3.8(-2.4)}$ for core(cusp) galaxies, respectively (Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]) combined with the significant dependence of $v_w$ on stochastic intermittency in the star formation history (Fig.~NICK).  The combination of (1) and (2) could well explain much of the scatter observed.

However, it is also possible that many of the galactic nuclei from the \citet{Miller+14} sample are in a thermally stable state.  The efficiency show in the right panel of Figure \ref{fig:miller} in some ways represents the maximum efficiency, since it does not take into account outflows from the accretion flow that would suppress the accreiton rate reaching large radii.  Nevertheless, many of the observed galaxies at low black hole masses of $M_{\bullet} \lesssim 10^{7}M_{\odot}$ lie well below the predicted average value for stable accretion (black line) and even approaching or above the maximally stable accretion rate $\dot{M}_{\rm TI}$ (yellow line) .  This suggests that these galaxies experience lower levels of heating than average galaxies of their mass, or that they are not accreting stably at all ($\S\ref{sec:stable}$).

@@

Figure \ref{fig:bh_growth} shows that across all black hole masses, Compton heating alone cannot prevent thermal instabilities ($\dot{M}_{\rm C} > \dot{M}_{\rm TI}$), as will described more below.  In the case of an old stellar population which produces insufficient heating to support steady accretion, this large scatter in Fig \ref{fig:Miller} can in principle also be understood by the wide range of gas masses experienced over the course of a cyclic episode between instabilities and periodic nuclear outbursts (e.g.~\citealt{Ciotti+10}).  The cyclic model also predicts that high $L_{X}$ nuclei should be more common in galaxies with younger stellar ages.  This appears in tension with the finding by \citet{Pellegrini10} that galaxies with younger stellar populations are observed to host less X-ray luminous nuclei, at least for stellar ages $<3$ Gyr.  This observation can also be understood in the stable accretion picture by the fact that young stars provide sufficient stellar wind heating to produce a thermally stable accretion state.

In general we find that core galaxies are more likely to possess higher accretion rates and potentially be thermally unstable.  This may be consistent with past studies which have found that core galaxies show a wide range of nuclear radio and X-ray luminosities, while cusp galaxies are confined below a threshold (e.g., \citealt{Bender+89, Pellegrini99, Capetti&Balmaverde05}).
